How to Avoid OOTA Without Really Trying

### 2024-10-11 paper/ootavacuous.tex
ootavacuous: Wordsmith last sentence of Section 6 lead-in
Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
ootavacuous: Yet more rework of Section 4.1's first sentence
Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

The exact definition of a computer language is a subject of some debate,
with standards, implementations, and users all having some degree of
influence~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2},
and each being prone to change over time.
In areas that are not well settled or where users might reasonably
want to resist the dictates of the standard,
compilers often provide switches to override their default behaviors.
An example is GCC's \co{-funsigned-char} command-line argument,
which causes it to treat variables of type \co{char} as unsigned
(see Section~\ref{sec:Semantic Dependencies and Architectural Variations}).
Many more examples of user control over language semantics
may be found through use of the command \co{make V=1}
in a Linux-kernel source tree and by the discussion
in~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2}.

@@

The definition given below is applicable only to C++ implementations
that treat all atomic objects as though they are volatile or
quasi volatile.
(For compilers that perform only per-thread analysis, not global
analysis, quasi volatile is sufficient.)
In this setting we can relate abstract and hardware executions by
means of the map of accesses described in
Section~\ref{sec:Volatile and Quasi Volatile Accesses}.
The key insight is that this allows us to consider semantic
dependencies at the level of the machine code, where they are much
more tractable.
This allows us to demonstrate that OOTA cycles cannot form
in the C++ implementations addressed by this paper.

>>>>

The exact definition of a computer language is a subject of some debate,
with standards, implementations, and users all having their own
influence~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2},
and each being prone to change over time.
In areas that are not well settled or where users might reasonably
want to resist the dictates of the standard,
compilers often provide switches to override their default behaviors.
An example is GCC's \co{-funsigned-char} command-line argument,
which causes it to treat variables of type \co{char} as unsigned
(see Section~\ref{sec:Semantic Dependencies and Architectural Variations}).
Many more examples of user control over language semantics
may be found through use of the command \co{make V=1}
in a Linux-kernel source tree and by the discussion
in~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2}.

@@

The definition given below is applicable only to C++ implementations
that treat all atomic objects as though they are volatile or
quasi volatile.
(For compilers that perform only per-thread analysis, not global
analysis, quasi volatile is sufficient.)
In this setting we can relate abstract and hardware executions by
means of the map of accesses described in
Section~\ref{sec:Volatile and Quasi Volatile Accesses}.
The key insight is that this allows us to consider semantic
dependencies at the level of the machine code, where they are much
more tractable.
Using this approach, we demonstrate that OOTA cycles cannot form in the
C++ implementations addressed by this paper.
